---
title: "Lesson Scripts"
author: "Victor Wildner"
date: "4/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Tips

Number as Integer:

```{r}
1L ## L Sufix
```

Access Atributes:

```{r}
x <- "hello"
attributes(x)
```

Set vector:

```{r}

```

Concatenate:

```{r}
a <- c("a", "b", "c")
a
```

Convert to something:

```{r}
x <- 0:6
as.numeric(x)
as.logical(x)
as.character(x)
```

Lists:

```{r}
x <- list(1, "a", TRUE, 1 + 4i)
x
```

Matrices:

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
m
dim(m)
```

Matrices:

```{r}
m <- 1:10
dim(m) <-c(2, 5)
m
```

cbiding and rbiding:

```{r}
x <- 1:3
y <- 10:12
cbind(x, y)
rbind(x, y)
```

Factors:

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x

table(x)

unclass(x)

```

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),
 levels = c("yes", "no"))
x
```

Missing Values:

```{r}
x <- c(1, 2, NA, 10, 3)
is.na(x)
is.nan(x)

y <- c(1, 2, NaN, NA, 4)
is.na(y)
is.nan(y)
```

Data Frames:

```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F)) 
x
nrow(x)
ncol(x)
```

Data Types - Name Attributes:

```{r}
x <- 1:3
names(x)
names(x) <- c("foo", "bar", "norf")
x

y <- list(a = 1, b = 2, c = 3)
y
m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("a", "b"), c("c", "d"))
m
```

## Reading Data Functions:

* **read.table**, **read.csv**, for reading tabular data
* **readLines**, for reading lines of a text file
* **source**, for reading in R code files (inverse of dump)
* **dget**, for reading in R code files (inverse of dput)
* **load**, for reading in saved workspaces
* **unserialize**, for reading single R objects in binary form

## Writing Data:

* **write.table**
* **writeLines**
* **dump**
* **dput**
* **save**
* **serialize**

## Reading Data read.table:

* **file**, the name of a file, or a connection
* **header**, logical indicating if the file has a header line
* **sep**, a string indicating how the columns are separated
* **colClasses**, a character vector indicating the class of each column in the dataset
* **nrows**, the number of rows in the dataset
* **comment.char**, a character string indicating the comment character
* **skip**, the number of lines to skip from the beginning
* **stringsAsFactors**, should character variables be coded as factors? (default=TRUE)

read.table
```{r}
## data <- read.table("foo.txt")
```
read.csv
```{r}
## data <- read.csv("foo.csv")
```

## Read Large Datasets

```{r}
## comment.char = ""
## initial <- read.table("datatable.txt", nrows = 100)
## classes <- sapply(initial, class)
## tabAll <- read.table("datatable.txt", colClasses = Classes)
```

## Estimate Memory Requirements

Supose you have 1.500.000 rows and 120 cols:
1.500.00 X 120 X 8 bytes/numeric = 1440000000 bytes
1440000000 bytes/2^(20) bytes/MB = 1.373.29 MB ~= 1.34 GB

## Deconstruct R Objects
dput-ting R Objects dput-ting R Objects (one object at a time)

```{r}
y <- data.frame(a = 1, b = "a")
dput(y)

# Put into a file
dput(y, file="y.R")
new.y <- dget("y.R")
new.y
```

dumping R Objects (various objects)

```{r}
x <- "foo"
y <- data.frame(a = 1, b = "a")
dump(c("x", "y"), file = "data.R")
rm(x, y)
source("data.R")
y
x
```

## Setting a Connection

```{r}
con <- file("y.R")
```

## Removing NA Values

```{r}
x <- c(1, 2, NA, 4, NA, 5)
bad <- is.na(x) # searches for missing elements
bad
x[!bad]

y <- c("a", "b", NA, "d", NA, "f")
good <- complete.cases(x, y) # searches for non missing vectors
good
x[good]
y[good]
```

